---
title: "purrr for (R)odeo"
author: "Ethan Addicott"
date: "11/5/2019"
output: 
  revealjs::revealjs_presentation:
    theme: night
    incremental: true
    transition: none
    css: slides.css
  beamer_presentation:
    colortheme: rose
    theme: Szeged
    toc: yes
---
# Slides and RMarkdown File
 http://github.com/eaddicott/library 

# Motivating purrr
<section style="text-align: left;">
*Iteration* is one of the most useful tasks a computer can perform for a researcher.

- Repetition
  - Static
  - Dynamic

- The programming *Rule of 3*
  - Never copy and paste more than twice
</section>

# Enter purrR
<section style="text-align: left;">

 We loop over vectors and dataframes, do some manipulation of the data, and save the results somewhere so often that there is a whole package designed to help us with this common task:

> <center> purrR </center>

 Thankfully this package is a part of the tidyverse, so we don't need to invoke it separately if we're already in the tidyverse library.
</section>

# Exercise: Summary Statistics
<section style="text-align: left;">
Your Task: Using your favorite R data (I'm using mtcars), generate column-by-column summary statistics
</section>
```{r}
head(mtcars)
```
## An Aside

Let's take the mean of a vector of numbers from 1 to 5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(microbenchmark)
```
```{r mtcars, echo=TRUE, paged.print=FALSE}
# The mean of a list of numbers
mean(1:5)
```

> This checks out! Can we do the same thing to generate the means by column for our data?

## FAIL! 
```{r}
#mean(mtcars)

#means(mtcars)
```
 These won't even run. Not a thing.

# Attempt 2
```{r , echo=TRUE, paged.print=FALSE}
output <- summary(mtcars)
output
```
## Now save the means
```{r}
means <- output[4,]
means
```

> Gross! I don't want to have to pull out a vector of doubles from this!

# Attempt 3: Super Common Issue
```{r, echo=TRUE}
#Preallocate the vector for means
means <- vector("double", ncol(mtcars))
# for loop to take the column by column means
for (i in length(mtcars)){
  means[[i]] <- mean(mtcars[[i]])
}
# see the mean vector
means
colnames(mtcars)
```
> What happened???

## Attempt 3: Redux
```{r, echo=TRUE}
#Preallocate
means <- vector("double", ncol(mtcars))
#for loop
for (i in seq_along(mtcars)){
  means[[i]] <- mean(mtcars[[i]])
}
means
```
Better...

> NOTICE seq_along solves the issue much better than length(mtcars[1,])

# purrr Solution
```{r, echo=TRUE}
#library(tidyverse)
means <- vector("double", ncol(mtcars))
means <- map(mtcars,mean)
means
```

## Getting the values
```{r}
means[1]
means[[1]]
25-means[[1]]
```

So we can extract the means and play with them.
> Still not convinced we should use purrr?
> It does accomplish things in fewer lines of code, but what else?

# Survey says...
```{r, echo=TRUE}
#library(microbenchmark)
mbm <- microbenchmark(
  "loop" = {means <- vector("double", ncol(mtcars))
            for (i in seq_along(mtcars)){
              means[[i]] <- mean(mtcars[[i]])
            }
  },
  "purrR" = {means <- vector("double", ncol(mtcars))
             means <- map(mtcars,mean)
             means <- as.numeric(means)
})
mbm
```
> purrr wins! 
> By over an order of magnitude!!


# The Details
- purrr functions run in C (read: fast!)
- map() writes to a vector
- There is a family of map functions that can write to a vector rather than a list (faster!)
  - map_dbl()
  - map_chr()
  - map_dfc(), map_dfr()
  - map_lgl()
  - map_int()
  
# Example: map_dbl()
```{r}
microbenchmark("map" = {means <- vector("double", ncol(mtcars))
                        means <- map(mtcars,mean)
                        means <- as.numeric(means)},
        "map_dbl()" = {means <- vector("double", ncol(mtcars))
                        means <- map_dbl(mtcars,mean)})
```
> Only slightly faster, but those microseconds can add up!

# Nested Dataframes
A short dive into the tidyverse here to explore some additional benefits in the purrr package.
```{r}
head(mtcars)
```
Let's group the data by the number of cylinders and create a nested dataframe.

## Nest by Cylinders
```{r}
nested <- mtcars %>%
  group_by(cyl) %>%
  nest()
nested
```

## This is reversible!
```{r}
nested %>% unnest()
```

# Models with Nests
Now we'll run a model (or set of models) over our nested dataframe.

```{r}
#Regressing weight on fuel economy
model_fn <- function(df){
  lm(mpg ~ wt,data = df)
}
#apply model over nested data and save to new model column
 m_df <- nested %>%
  mutate(model = map(data,model_fn))
#take a look
 m_df
```
> Let's view the results!

## Viewing the results
 
```{r}
 results <- m_df$model %>% map(summary) 
 results
```

## Extract R squared
```{r}
 results %>% map_dbl("r.squared")
```

## Extract Coefficients
```{r}
 results %>% map("coefficients")
```

## Extract Coefficients
```{r}
 betas <- function(model){
   coefficients(model)[[2]]
 }
 
 results %>% map_dbl(betas)
 
```

# map2: Another feature in purrr

- Sick of figuring out how to use an apply function over two objects?
- mapply got you down?
- FEAR NOT!

## Using map2
```{r}
mtcars %>% mutate(hp_wt_ratio = map2_dbl(hp,wt, ~ .x / .y)) %>% select(hp_wt_ratio)
```
## Using map2
```{r}
mtcars %>% mutate(hp_wt_ratio = map2_dbl(hp,wt,`/`)) %>% select(hp_wt_ratio)
```
# Resources

- https://r4ds.had.co.nz/iteration.html
- purrR Cheatsheet: https://github.com/rstudio/cheatsheets/blob/master/purrr.pdf
- https://github.com/cwickham/purrr-tutorial
- https://emoriebeck.github.io/R-tutorials/purrr/


