---
title: "purrr for (R)odeo"
author: "Ethan Addicott"
date: "11/5/2019"
output:
  beamer_presentation:
    colortheme: rose
    theme: Szeged
    toc: yes
  slidy_presentation: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(microbenchmark)
```
## Motivating purrr
*Iteration* is one of the most useful tasks a computer can perform for a researcher.

- Repetition
  - Static
  - Dynamic

- The programming *Rule of 3*
  - Never copy and paste more than twice

## Enter purrR

We loop over vectors and dataframes, do some manipulation of the data, and save the results somewhere so often that there is a whole package designed to help us with this common task: purrR

Thankfully this package is a part of the tidyverse, so we don't need to invoke it separately if we're already in the tidyverse library.

## Exercise: Column by Column Summary Statistics
```{r mtcars, echo=TRUE, paged.print=FALSE}
#mean(mtcars) 
#means(mtcars)
```
These won't even run. Not a thing.

## Attempt 2
```{r , echo=TRUE, paged.print=FALSE}
output <- summary(mtcars)
means <- output[4,]
means
```

Gross

## Attempt 3: Super Common Issue
```{r, echo=TRUE}
means <- vector("double", ncol(mtcars))
for (i in length(mtcars)){
  means[[i]] <- mean(mtcars[[i]])
}
means
```
## Attempt 3: Redux
```{r, echo=TRUE}
means <- vector("double", ncol(mtcars))
for (i in seq_along(mtcars)){
  means[[i]] <- mean(mtcars[[i]])
}
means
```
Better...

## purrr Solution
```{r, echo=TRUE}
#library(tidyverse)
means <- vector("double", ncol(mtcars))
means <- map(mtcars,mean)
means
```
## Survey says...
```{r, echo=TRUE}
#library(microbenchmark)
mbm <- microbenchmark(
  "loop" = {means <- vector("double", ncol(mtcars))
            for (i in seq_along(mtcars)){
              means[[i]] <- mean(mtcars[[i]])
            }
  },
  "purrR" = {means <- vector("double", ncol(mtcars))
             means <- map(mtcars,mean)
             means <- as.numeric(means)
})
mbm
```
purrr wins! By over an order of magnitude.


## The Details
- purrr functions run in C (read: fast!)
- map() writes to a vector
- There is a family of map functions that can write to a vector rather than a list (faster!)
  - map_dbl()
  - map_chr()
  - map_dfc(), map_dfr()
  - map_lgl()
  - map_int()
  
## Example: map_dbl()
```{r}
microbenchmark("map" = {means <- vector("double", ncol(mtcars))
                        means <- map(mtcars,mean)
                        means <- as.numeric(means)},
        "map_dbl()" = {means <- vector("double", ncol(mtcars))
                        means <- map_dbl(mtcars,mean)})
```
## Nested Dataframes
```{r}
head(mtcars)
nested <- mtcars %>%
  group_by(cyl) %>%
  nest()
nested
# This is reversible!
nested %>% unnest()
```
## Models with Nests
```{r}
model_fn <- function(df){
  lm(mpg ~ wt,data = df)
}

 m_df <- nested %>%
  mutate(model = map(data,model_fn))
 
 m_df
```
## View the results
 
```{r}
 results <- m_df$model %>% map(summary) 
 results
 results %>% map_dbl("r.squared")
 
 results %>% map("coefficients")
 
 betas <- function(model){
   coefficients(model)[[2]]
 }
 
 results %>% map_dbl(betas)
 
```

## Also map2

```{r}
mtcars %>% mutate(hp_wt_ratio = map2_dbl(hp,wt, ~ .x / .y)) %>% select(hp_wt_ratio)
mtcars %>% mutate(hp_wt_ratio = map2_dbl(hp,wt,`/`)) %>% select(hp_wt_ratio)

```
## Resources

- https://r4ds.had.co.nz/iteration.html
- purrR Cheatsheet: https://github.com/rstudio/cheatsheets/blob/master/purrr.pdf
- https://github.com/cwickham/purrr-tutorial
- https://emoriebeck.github.io/R-tutorials/purrr/


